---
title: "Homework 3"
author: Eman Ibrahim (ei2291)
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1

Import the instacart data

```{r}
data("instacart")
```

The dataset tells us briefly about how many people ordered their groceries online through instacart. It tells us what they ordered etc. The dataset has 1,384,617 observations and 15 variables.Since each row in this dataset shows a product with a user's order, Some key variables are `user_id`. `product_id`, `order_id`, `order_number`, `aisle`, `product_name`, `order_number`.

```{r}
head(instacart,5)
```

This gives us the first five rows.

```{r}
total_aisles_df = 
  instacart |> 
  summarize(total_aisles_df=n_distinct(aisle))

most_aisle_counts =
  instacart |> 
  group_by(aisle_id, aisle) |> 
  summarize(n_items=n()) |> 
  arrange(desc(n_items))
```
The total number of aisle is 134.
The top five aisles that have the most items ordered from are aisles 83, 24, 123,120, 21 with fresh vegetables, fresh fruits, packed vegetables fruits, yogurt then packaged chips, respectively.  


Plot for number of items ordered in each aisle >10,000
```{r}
instacart |> 
  group_by(aisle) |> 
  summarize(n_items=n()) |> 
  filter(n_items> 10000) |> 
  arrange(desc(n_items)) |>
  ggplot(aes(x=reorder(aisle, n_items), y=n_items))+
  geom_col( fill ="steelblue")+
    labs(
    x= "Aisle",
    y = "No. of items ordered",
    title = "No. of items ordered in each Aisle for >10,000"
  )+
  coord_flip()+
  theme_minimal()
```

Table showing top 3 items in the selected aisles:
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(n_orders=n()) |>
  slice_max(n_orders, n=3) |>
  arrange(aisle, desc(n_orders)) |> 
  knitr::kable()
  
```


Mean hour of the day for Pink Lady Apples and Coffee Ice Cream by days of the week:
```{r}
instacart |> 
  select (product_name, order_hour_of_day, order_dow) |> 
  filter (product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |> 
  summarize(mean_hour=round (mean(order_hour_of_day, na.rm = TRUE), 2)
  ) |> 
  mutate(
    order_dow = recode(order_dow, 
    `0` = "Sunday",
    `1` = "Monday",
    `2` = "Tuesday",
    `3` = "Wednesday",
    `4` = "Thursday",
    `5` = "Friday",
    `6` = "Saturday"
                     )
  ) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )|> 
  knitr::kable()
```

## Problem 2

Import and clean dataset

```{r}
zipcodes_county_df=
  read_csv("Zip Codes.csv") |> 
  janitor::clean_names() |> 
  mutate(
    county = case_when(
      county == "New York" ~ "Manhattan",
      county == "Kings" ~ "Brooklyn",
      county == "Richmond" ~ "Staten Island",
      TRUE ~ county
    )
  ) |> 
  distinct(zip_code,.keep_all=TRUE)
```

Import and clean dataset 2

```{r}
zipcodes_neighborhood_df=
  read_csv("Zip codes neighborhood.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = starts_with ("x20"),
    names_to= "date",
    values_to = "zori"
  ) |> 
  rename(county=county_name,
         zip_code=region_id) |> 
  mutate(date = lubridate::ymd(str_remove(date, "^x")),
         county = case_when(
    county == "New York County" ~ "Manhattan",
    county == "Kings County" ~ "Brooklyn",
    county == "Richmond County" ~ "Staten Island",
    county == "Queens County" ~ "Queens",
    county == "Bronx County" ~ "Bronx",
    TRUE ~ county)) 
```

Join datasets
```{r}
combined_df=
  left_join(zipcodes_neighborhood_df, zipcodes_county_df, by="zip_code") |> 
  select(-county.y) |> 
  rename(county = county.x) |> 
  relocate(zip_code,county) |> 
  arrange(zip_code, county)
```

```{r}
combined_df |> 
  group_by(zip_code) |> 
  summarize(zip_observed = n()) |> 
  filter(zip_observed == 116) |> 
  nrow()
```

```{r}
combined_df |> 
  group_by(zip_code) |> 
  summarize(zip_observed = n()) |> 
  filter(zip_observed < 10) 
```

149 zip codes were obsereved 116 times, and 0 zipcodes were observed less than 10 times from January 2015 to August 2024.Some zipcodes are observed less than other because either data might be missing for certain months or those areas have few rental listings, while for those observed in each month are probably ones located in active areas and have data throughout the period.

Average rental price in each borough and year

```{r}
combined_df |> 
  select(date, county, zori) |> 
  mutate(year = year(date)) |> 
  group_by(county, year) |> 
  summarize(avg_rent = mean(zori, na.rm =TRUE) ) |> 
  arrange (county, year, avg_rent) |> 
  pivot_wider(
    names_from = year,
    values_from = avg_rent
  ) |> 
  knitr::kable(digits=2)
  
```

Throughout all the boroughs, the average rental prices increased from 2015 to 2024. Staten Island initially didn't have data and first data input was in 2020. Manhattan had the highest prices throughout, followed by Brooklyn and Queens.


```{r}
combined_df |> 
  select(date, county, zori, zip_code) |> 
  mutate(year = year(date)) |> 
  ggplot(aes(x=year, y=zori, group=zip_code, color=county))+
  geom_line(alpha=0.3)+
  stat_summary(aes(group=county),fun = median, geom ="line",  color = "black" )+
  facet_wrap(~county, scales = "free_y") +
  scale_x_continuous(
    breaks=seq(2015, 2024, by = 1),
    expand = c(0.01, 0)
    ) +
   labs(
    title = "NYC Rental Prices by ZIP Code Across Boroughs (2015-2024)",
    x = "Year",
    y = "ZORI (rent)",
    color = "borough")+
  theme_minimal()
  
```

This is a plot of the different zipcode level rent price trends for the five boroughs from 2015 to 2024. As we can see in the graph, rental prices overall increased in all the borough with a slight dip in 2020 which is likely due to covid-19. Staten Island data later became available starting 2020 and it generally has lower rent. Manhattan has he highest rent followed by Brooklyn and Queens.


```{r}
combined_df |> 
  filter(year(date) ==2023) |> 
  group_by(zip_code, county, date) |> 
  summarize(avg_rent = mean(zori, na.rm=TRUE), .groups="drop") |> 
  ggplot(aes(x=county, y=avg_rent, fill=county))+
  geom_boxplot(alpha=0.7)+
  labs(
    title = "Zipcode levels for averagre rent prices in 2023",
    x = "Borough",
    y = "Average rent price",
    fill = "Borough"
         )+
  theme_minimal()
```
Based on the plot, we can see that Manhattan had the highest rent price then Brooklyn then Queens then Bronx. Staten Island had the lowest average rent. Manhattan then Brooklyn show variation of prices based on zipcode.

```{r}
library (patchwork)
```

```{r}
plot_year = combined_df |> 
  select(date, county, zori, zip_code) |> 
  mutate(year = year(date)) |> 
  ggplot(aes(x=year, y=zori, group=zip_code, color=county))+
  geom_line(alpha=0.3)+
  stat_summary(aes(group=county),fun = median, geom ="line",  color = "black" )+
  facet_wrap(~county, scales = "free_y") +
  scale_x_continuous(
    breaks=seq(2015, 2024, by = 1),
    expand = c(0.01, 0)
    ) +
   labs(
    title = "NYC Rental Prices by ZIP Code Across Boroughs (2015-2024)",
    x = "Year",
    y = "ZORI (rent)",
    color = "borough")+
  theme_minimal()

plot_2023=combined_df |> 
  filter(year(date) ==2023) |> 
  group_by(zip_code, county, date) |> 
  summarize(avg_rent = mean(zori, na.rm=TRUE), .groups="drop") |> 
  ggplot(aes(x=county, y=avg_rent, fill=county))+
  geom_boxplot(alpha=0.7)+
  labs(
    title = "Zipcode levels for averagre rent prices in 2023",
    x = "Borough",
    y = "Average rent price",
    fill = "Borough"
         )+
  theme_minimal()

combined_plot = plot_year/plot_2023
```

```{r}
ggsave(
  filename="results/nyc_combined_rentals.png",
  plot = combined_plot,
  width=12,
  height =10,
  bg="white"
)
```

## Problem 3

load and clean 
```{r}
nhanes_demo =
  read_csv("nhanes_covar.csv", skip=4) |> 
  janitor::clean_names() |> 
  mutate(
    sex = recode(sex,
         `1` = "male",
         `2` = "female"
                 ),
    education = recode(education,
         `1` = "less than high school",
         `2` = "high school equivalent",
         `3` = "more than high school"
                       )
  ) |> 
filter(age >=21) |> 
  drop_na()
```

```{r}
nhanes_accel=
  read_csv("nhanes_accel.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = starts_with("min"),
    names_to = "minute",
    values_to = "MIMS"
  ) |> 
  mutate(minute=as.integer(str_remove(minute, "min"))) |> 
drop_na()
```

merging both

```{r}
nhanes_combined = 
  nhanes_accel |> 
  inner_join(nhanes_demo, by ="seqn")
```

table for number of men and women in each education category

```{r}
nhanes_combined |> 
  group_by(education, sex) |> 
  summarize(n = n_distinct(seqn), .groups = "drop") |> 
  pivot_wider(
    names_from = sex,
    values_from = n
  ) |> 
  knitr::kable()
```
From the table the most participants are with more than high school education. In those female are the most closely followed by male. for the less than high school there are almost the same number of participants from both sex. the greatest difference is seen in the high school equivalent, where male outnumber female significantly.


Visualization of the age distribution for men and women in each category

```{r}
nhanes_combined |> 
  select(sex, education, age, seqn) |> 
  distinct() |> 
  ggplot(aes(x=age, fill=sex))+
  geom_histogram(position = "dodge",binwidth = 5, color = "black") + 
  facet_wrap(~ education) +
  labs(
    title = "Age Distribution by Sex and Education",
    x = "Age",
    y = "The number of participants",
    fill = "Sex"
  )+
  theme_minimal()
```
Based on the plot the most participants are people with more than high school education with female age group from 20-40 being the most.

Total Daily Activity Vs. Age by Sex and Education

```{r} 
nhanes_total =
  nhanes_combined |> 
  group_by(seqn, sex, age, education) |>   summarize(tot_MIMS = round(sum(MIMS, na.rm=TRUE), 2), .group="drop") 

nhanes_total |> 
  ggplot(aes(x=age, y=tot_MIMS, color =sex))+
  geom_point(alpha=0.6) +
  geom_smooth(SE = FALSE, method = "loess")+
  facet_wrap(~education)+
  labs(
    title = "Total daily activity vs age by sex and education",
    x = "Age",
    y = "Total MIMS",
    color = "Sex"
  )+
  theme_minimal()

```

As seen in the plot, in individuals with a high school equivalent, females had the highest average total MIMS (total daily activity) especially those from age group 30 to 50. In the less than high school education category. the peak is seen around age group 60, where male had the overall average higher average total daily activity. For individuals with more than high school education, the trendline was pretty constant but on average female through out had a higher average total daily activity, but both sex were more or less close. Something seen in both sex and education level is that as age increases with 60+ average total daily activity gradually decreases.

24-Hour Activity by Sex and Education Level

```{r}
nhanes_minute =
  nhanes_combined |> 
  group_by(minute, sex, education) |> 
  summarize(mean_MIMS = round(mean(MIMS, na.rm = TRUE), 2), .group="drop")

nhanes_minute |> 
  ggplot(aes(x=minute, y= mean_MIMS, color =sex)) +
  geom_line (alpha=0.8)+
  geom_smooth( se = FALSE, method = "loess")+
  facet_wrap(~education, ncol=1)+
  labs(
    title = "24 hour Activity by Sex and Education",
    x = "Minute of the day",
    y = "Average MIMS/Activity",
    color = "Sex"
  )+
  theme_minimal()
```
In all three education level plots, we see the highest average activity to be from the 500 to about 1000 minute of the day which then gradually declines. Both male and female have about the same/similar average activity by minute of the day especially for those that have a less than high school equivalent and high school equivalent. For those that have more than a high school education, we see that female for the most part throughout the day had a higher average MIMS.






















