---
title: "Homework 3"
author: Eman Ibrahim (ei2291)
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1

Import the instacart data

```{r}
data("instacart")
```

The dataset tells us briefly about how many people ordered their groceries online through instacart. It tells us what they ordered etc. The dataset has 1,384,617 observations and 15 variables.Since each row in this dataset shows a product with a user's order, Some key variables are `user_id`. `product_id`, `order_id`, `order_number`, `aisle`, `product_name`, `order_number`.

```{r}
head(instacart,5)
```

This gives us the first five rows.

```{r}
total_aisles_df = 
  instacart |> 
  summarize(total_aisles_df=n_distinct(aisle))

most_aisle_counts =
  instacart |> 
  group_by(aisle_id, aisle) |> 
  summarize(n_items=n()) |> 
  arrange(desc(n_items))
```
The total number of aisle is 134.
The top five aisles that have the most items ordered from are aisles 83, 24, 123,120, 21 with fresh vegetables, fresh fruits, packed vegetables fruits, yogurt then packaged chips, respectively.  


Plot for number of items ordered in each aisle >10,000
```{r}
instacart |> 
  group_by(aisle) |> 
  summarize(n_items=n()) |> 
  filter(n_items> 10000) |> 
  arrange(desc(n_items)) |>
  ggplot(aes(x=reorder(aisle, n_items), y=n_items))+
  geom_col( fill ="steelblue")+
    labs(
    x= "Aisle",
    y = "No. of items ordered",
    title = "No. of items ordered in each Aisle for >10,000"
  )+
  coord_flip()+
  theme_minimal()
```

Table showing top 3 items in the selected aisles:
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(n_orders=n()) |>
  slice_max(n_orders, n=3) |>
  arrange(aisle, desc(n_orders)) |> 
  knitr::kable()
  
```


Mean hour of the day for Pink Lady Apples and Coffee Ice Cream by days of the week:
```{r}
instacart |> 
  select (product_name, order_hour_of_day, order_dow) |> 
  filter (product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |> 
  summarize(mean_hour=round (mean(order_hour_of_day, na.rm = TRUE), 2)
  ) |> 
  mutate(
    order_dow = recode(order_dow, 
    `0` = "Sunday",
    `1` = "Monday",
    `2` = "Tuesday",
    `3` = "Wednesday",
    `4` = "Thursday",
    `5` = "Friday",
    `6` = "Saturday"
                     )
  ) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )|> 
  knitr::kable()
```












